---
title: "A3: Incarceration"
output: html_document
---

### Introduction

An introduction of the problem domain and a description of the variable(s) you are choosing to analyze (and why!)
In the US the discrimination and unfair policing of people of color has impacted the prison system since it's beginning. This discrimination in both the police and court systems has resulted in increased jailing of people of color. I am choosing to look at variables from the data set pertaining to how many prisoners there are of different races. This includes the total populations of black, latinx, and white people in the prisons.  
  

### Summary Information

The data shows a clear discrepancy in the number of imprisonments according to the race of a person. Across all of the data through the years, I calculated that in a prison almost 40% of the inmates are black, about 13% are latinx, and roughly 31% are white. When considering that a majority of the US is white, and African Americans only make up 14% of the population, a clear injustice is displayed. This injustice is also displayed in the percent of people in the US in prison, sorted by race. The data shows that 1.7% of black people in the US are in prison. This statistic is only 0.5% for latinx Americans, and 0.26% for white Americans. This shows that African Americans are over 6 times more likely to be put in prison than white Americans. 


### The Dataset

This data was collected by The Vera Institute. This group is working to combat the over-policing and mass incarceration of people in poverty and people of color in the US. This data was collected from government surveys of the jail system that are conducted by the US government every five to eight years. These surveys are conducted by the US Department of Justice Bureau of Justice Statistics. This data is collected every few years because in 1965 the President’s Commission on Law Enforcement and Administration of Justice requested a census of the jail system. These surveys have been performed since then. The data consists of 153811 rows and 36 columns. There are many ethical questions and questions of power that arise when working with this data. Data like this has been looked at many times to draw racist and biased conclusions. Some people look at the increased rates of incarceration among African Americans and use it to justify their racist worldviews. Instead of considering the nuanced and biased justice system, they assume that all of the people arrested should have been arrested, and ignore any outside factors. Every societal and systemic facet of this problem must be considered when looking at this data. This data could be limited due to how it is not collected very often- only every 5 to 8 years.


### Trends Over Time Chart

This graphs displays the change in the number of black, white and Latinx inmates over the years, beginning in 1983 and ending in 2016. This graph displays a clear disparity in incarceration between different racial groups when considering the much larger white population in the US. This disparity is shown to be improving over the years however as the number of black inmates decreased over the last couple decades, however people of color are still incarcerated at a much higher rate.

```{r, echo = FALSE, code = readLines("chart1_example.R")}
```

### Variable Comparison Chart

This chart shows the percentage of three different racial groups that were in prison at the time of the prison census in 2016. I included this chart to show the large disparity in incarceration between different races in the US. By putting the data into percentages of the 15 to 64 year old population instead of just the total numbers, you are able to see the injustices people of color face in the US.

```{r, echo = FALSE, code = readLines("chart2_example.R")}
```

### Map

This graph shows a clear disparity in the number of people in prison by state and region of the US. The states with the most people in prison seem to be the ones typically though of as poorer, and with histories of racism and discrimination. The state and region a person lives in can affect their life experience and experience with the US justice system.

```{r, echo = FALSE, code = readLines("map_example.R")}
```
